Tropical Mathematics and the Lambda Calculus

### 2023-01-24 appV.tex
Commit notturno

\begin{proposition}
All tropical functions $f: Q^{X}\to Q^{Y}$ are concave.
\end{proposition}
\begin{proof}
Let us first show that all functions of the form $f( x )_{b}= \mu  x + c$ are concave:
we have $f(\alpha x + (1-\alpha) y  )_{b}= \mu(\alpha x )+(1-\alpha) y  )+c=
 \mu(\alpha x )+(1-\alpha) y  )+\alpha c+(1-\alpha)c=
 \alpha(\mu  x  + c)+(1-\alpha)(\mu  y  +c)=\alpha f( x )_{b}+(1-\alpha) f( x )_{b}$.

@@

\begin{theorem}[Theorem \ref{theorem:fepsilon}]
Let $f:\overline{\R}_{\geq 0}^{\{1,\dots, K\}} \to \overline{\R}_{\geq 0}$ tropical analytical, with matrix $\hat f : \N^{K} \to \overline{\R}_{\geq 0}$.
Then, for all $0<\epsilon<+\infty$, there is $\mathcal{F}_\epsilon\subseteq\N^{K}$ s.t.\
\begin{enumerate}
 \item $\mathcal{F}_\epsilon$ is finite
 \item If $\mathcal{F}_\epsilon= \emptyset$ then $f( x ) = +\infty$ for all $ x \in \overline{\R}_{\geq 0}^{\{1,\dots, K\}}$
 \item If $f( x _0) = +\infty$ for some $ x _0\in [\epsilon,\infty)^{K}$ then $\mathcal{F}_\epsilon= \emptyset$
 \item The restriction of $f$ to $[\epsilon,+\infty]^{K}$ coincides with the  \emph{tropical polynomial}:
      \[
        P_\epsilon( x )= \min\limits_{ n  \in\mathcal{F}_\epsilon} \set{\hat f ( n  ) +  n   x }.
      \]
      where $ n   x = \sum_{i=1}^{K}n_{i}x_{i}$.
\end{enumerate}
\end{theorem}
\begin{proof}
We let $\mathcal F_\epsilon$ to be the complementary in $\N$ of the set:
\[
 \set{ n  \in\N^{K} \mid \textit{either } \hat f ( n  )=+\infty \textit{ or there is }  m  \prec  n  \textit{ s.t.\ } \hat f( m  )\leq\hat f( n  )+\epsilon}.
\]
In other words, $ n  \in\mathcal F_\epsilon$ iff $\hat f( n  )<+\infty$ and for all $ m  \prec  n  $, one has $\hat f( m  )>\hat f( n  )+\epsilon$.

@@

Due to the previous discussion about the impossibility of stating Corollary \ref{cor:cont} in the case where one of the coordinates of $x$ is $0$, the continuity of tropical functions on the hyperplanes $\mathcal H_a:=\set{x\in {\R}_{\geq 0}^X \mid x_a=0}$, for $a\in X$, must be treated separately.
Similarly, the continuity at points with infinite coordinates must be treated separately as well.
We left it for future investigations.

@@

 Remark also that tropical functions have no reason to be linear nor sublinear.
\end{remark}

\begin{remark}\label{rmk:tropMonot}
 Tropical functions are monotone w.r.t.\ to the cone order on its domain and codomain.
 This is clear, using the definition of tropical functions, because $\mu x\leq \mu y$ if $x\leq y$.
\end{remark}

@@

\begin{theorem}[Theorem~\ref{thm:ScottCont}]
 All tropical functions $f:\overline{\R}^X_{\geq 0}\to\overline{\R}^Y_{\geq 0}$ are Scott-continuous on $\R^X_{> 0}$ w.r.t.\ the cone-orders on its domain and codomain.
\end{theorem}
\begin{proof}
 Let $(x_i)_i$ a directed net in $\R^X_{> 0}$ s.t.\ $\bigvee\limits_i x^i$ exists in $\R^X_{> 0}$.
 Then $\inf\limits_i \supnorm{\bigvee\limits_i x^i - x^i} =0$, where $\bigvee\limits_i x^i - x^i$ exists because $\bigvee\limits_i x^i \geq x^i$ for all $i$.
 Since $f$ is $\supnorm{.}$-continuous on $\R^X_{> 0}$, then $\inf\limits_i \supnorm{f(\bigvee\limits_i x^i) - f(x^i)} =0$, where $f(\bigvee\limits_i x^i) - f(x^i)$ exists because $f(\bigvee\limits_i x^i) \geq f(x^i)$ for all $i$ being $f$ monotone (Remark \ref{rmk:tropMonot}).
 We can therefore apply Proposition \ref{prop:infsup} to the directed net $(f(x^i))_i$ in $\overline{\R}^Y_{\geq 0}$, obtaining that $\bigvee\limits_i f(x^i)$ exists in $\overline{\R}^Y_{\geq 0}$ and it coincides with $f(\bigvee\limits_i x^i)$.
\end{proof}

>>>>

\begin{proposition}
All tLs $f: Q^{X}\to Q^{Y}$ are concave.
\end{proposition}
\begin{proof}
Let us first show that all functions of the form $f( x )_{b}= \mu  x + c$ are concave:
we have $f(\alpha x + (1-\alpha) y  )_{b}= \mu(\alpha x )+(1-\alpha) y  )+c=
 \mu(\alpha x )+(1-\alpha) y  )+\alpha c+(1-\alpha)c=
 \alpha(\mu  x  + c)+(1-\alpha)(\mu  y  +c)=\alpha f( x )_{b}+(1-\alpha) f( x )_{b}$.

@@

\begin{theorem}[Theorem \ref{theorem:fepsilon}]
Let $k\in\N$ and $f:\Lawv^k\to\Lawv$ a tLs with matrix $\matr f:\N^k\to\Lawv$.
 For all $0<\epsilon<\infty$, there is $\C F_\epsilon \subseteq \N^k$ such that:
\begin{enumerate}
 \item $\C F_\epsilon$ is finite
 \item If $\mathcal{F}_\epsilon= \emptyset$ then $f( x ) = +\infty$ for all $ x \in \Lawv^k$
 \item If $f( x _0) = +\infty$ for some $ x _0\in [\epsilon,\infty)^{K}$ then $\mathcal{F}_\epsilon= \emptyset$
 \item The restriction of $f$ on $[\epsilon,\infty]^k$ coincides  with the tropical {polynomial} \[P_\epsilon(x):=\min\limits_{n\in \C F_\epsilon}\set{nx+\matr f(n)}\]
where $nx:=\sum_{i=1}^k n_ix_i$.
\end{enumerate}
\end{theorem}
\begin{proof}
We let $\mathcal F_\epsilon$ to be the complementary in $\N$ of the set:
\[
 \set{ n  \in\N^{K} \mid \textit{either } \hat f ( n  )=+\infty \textit{ or there is }  m  \prec  n  \textit{ s.t.\ } \hat f( m  )\leq\hat f( n  )+\epsilon}.
\]
In other words, $ n  \in\mathcal F_\epsilon$ iff $\hat f( n  )<+\infty$ and for all $ m  \prec  n  $, one has $\hat f( m  )>\hat f( n  )+\epsilon$.

@@

Due to the previous discussion about the impossibility of stating Corollary \ref{cor:cont} in the case where one of the coordinates of $x$ is $0$, the continuity of tLs on the hyperplanes $\mathcal H_a:=\set{x\in {\R}_{\geq 0}^X \mid x_a=0}$, for $a\in X$, must be treated separately.
Similarly, the continuity at points with infinite coordinates must be treated separately as well.
We left it for future investigations.

@@

 Remark also that tLs have no reason to be linear nor sublinear.
\end{remark}

\begin{remark}\label{rmk:tropMonot}
 tLs are monotone w.r.t.\ to the cone order on its domain and codomain.
 This is clear, using the definition of tLs, because $\mu x\leq \mu y$ if $x\leq y$.
\end{remark}

@@

\begin{theorem}[Theorem~\ref{thm:ScottCont}]
 All tLs $f:\overline{\R}^X_{\geq 0}\to\overline{\R}^Y_{\geq 0}$ are Scott-continuous on $\R^X_{> 0}$ w.r.t.\ the cone-orders on its domain and codomain.
\end{theorem}
\begin{proof}
 Let $(x_i)_i$ a directed net in $\R^X_{> 0}$ s.t.\ $\bigvee\limits_i x^i$ exists in $\R^X_{> 0}$.
 Then $\inf\limits_i \supnorm{\bigvee\limits_i x^i - x^i} =0$, where $\bigvee\limits_i x^i - x^i$ exists because $\bigvee\limits_i x^i \geq x^i$ for all $i$.
 Since $f$ is $\supnorm{.}$-continuous on $\R^X_{> 0}$, then $\inf\limits_i \supnorm{f(\bigvee\limits_i x^i) - f(x^i)} =0$, where $f(\bigvee\limits_i x^i) - f(x^i)$ exists because $f(\bigvee\limits_i x^i) \geq f(x^i)$ for all $i$ being $f$ monotone (Remark \ref{rmk:tropMonot}).
 We can therefore apply Proposition \ref{prop:infsup} to the directed net $(f(x^i))_i$ in $\overline{\R}^Y_{\geq 0}$, obtaining that $\bigvee\limits_i f(x^i)$ exists in $\overline{\R}^Y_{\geq 0}$ and it coincides with $f(\bigvee\limits_i x^i)$.
\end{proof}

### 2023-01-24 section3C.tex
Commit notturno

\begin{corollary}\label{cor:T(M)=M}
 The interpretation of the $\STDLC$-Taylor expansion of a $\STLC$-term $M$ given in \eqref{eq:taylor}, converges to the interpretation of $M$.
\end{corollary}

>>>>

\begin{corollary}\label{cor:T(M)=M}
 In $\LREL_!$, the interpretation of the Taylor expansion of a $\STLC$-term $M$, given in \eqref{eq:taylor}, converges to the one of $M$.
\end{corollary}
